---
title: "Untitled"
author: "Anina"
date: "4 6 2020"
output: pdf_document
---


Load Packages 


```{r}
install.packages("readr")
library(readr)
library(tidyverse)
library(dplyr)

```


Poverty Rate. 2013-2018

```{r setup, include=FALSE}
poverty <- read.csv("Poverty_Status_ACS5_2018.csv", header=TRUE)
View(poverty)

poverty <- poverty %>%
  select(NAME, S1701_C03_001E) %>%
  rename(pct_below_poverty_level = S1701_C03_001E, state=NAME) %>%
  slice(-c(1))

poverty$pct_below_poverty_level <- as.numeric(levels(poverty$pct_below_poverty_level))[poverty$pct_below_poverty_level]
str(poverty)

```



Employment Status & Income. 2013-2018

```{r setup, include=FALSE}
empl_HHincome <- read.csv("EmploymentStatus_HHIncome_ACS5_2018.csv", header=TRUE)
View(empl_HHincome)

empl_HHincome <- empl_HHincome %>%
  select(NAME, DP03_0009PE, DP03_0063E) %>%
  rename(state=NAME, unemplyment_rate = DP03_0009PE, mean_HH_income = DP03_0063E) %>%
  slice(-c(1))

str(empl_HHincome)

empl_HHincome$unemplyment_rate <- as.numeric(levels(empl_HHincome$unemplyment_rate))[empl_HHincome$unemplyment_rate]
empl_HHincome$mean_HH_income <- as.numeric(levels(empl_HHincome$mean_HH_income))[empl_HHincome$mean_HH_income]

```



Educational Attainment, 2013-2018

```{r setup, include=FALSE}
educ <- read.csv("Education_Language_ACS_2018.csv", header=TRUE)
View(educ)

### Data Inspection
# # Look for variables
# M_edu <- as.matrix(educ)
# Row1_edu <- M_edu[1,]
# str(Row1_edu)
# 
# ### Find Columns that we need ###
# 
# Row1_edu[grep("EDUCATIONAL ATTAINMENT", Row1_edu)]
# # Row1_edu[grep("Less than 9th grade", Row1_edu)]
# # # DP02_0059E 
# # # DP02_0059PE 
# # 
# # Row1_edu[grep("9th to 12th grade", Row1_edu)]
# # # DP02_0060E 
# # # DP02_0060PE 
# # 
# # Row1_edu[grep("High school graduate", Row1_edu)]
# # # DP02_0061E  
# # # DP02_0061PE 
# # 
# # Row1_edu[grep("Some college, no degree", Row1_edu)]
# # # DP02_0062E
# # # DP02_0062PE
# # 
# # Row1_edu[grep("Associate's degree", Row1_edu)]
# # # DP02_0063E
# # # DP02_0063PE 
# # 
# # Row1_edu[grep("Bachelor's degree", Row1_edu)]
# # # DP02_0064E
# # # DP02_0064PE 
# # 
# # Row1_edu[grep("Graduate or professional degree", Row1_edu)]
# # # DP02_0065E
# # # DP02_0065PE 
# 
# 
# grep("Less than 9th grade", Row1_edu)
# grep("Graduate or professional", Row1_edu)
# # 235 - 262

educ <- educ %>%
  select(NAME, c(235:262)) 


educ <- educ %>%
  select(NAME, ends_with("E"), ends_with("PE"))

View(educ)
str(educ)

# Change variables from factor to numeric

my_function <-function(x) {
  as.numeric(levels(x))[x]}

educ <- educ %>%
  mutate(NAME = as.character(NAME)) %>%
  mutate_if(is.factor, my_function )
str(educ)
View(educ)


educ <- educ %>%
  mutate( tot_less_than_high_school = rowSums(educ[, c("DP02_0059E", "DP02_0060E")], na.rm = TRUE)) %>% # taking together people with <9th grade and 9th to 12th grade
  mutate( pct_less_than_high_school = rowSums(educ[, c("DP02_0059PE", "DP02_0060PE")], na.rm = TRUE)) %>%
  mutate( tot_college_associate = rowSums(educ[, c("DP02_0062E", "DP02_0063E")], na.rm = TRUE)) %>% # group college and associate degree together
  mutate( pct_college_associate = rowSums(educ[, c("DP02_0062PE", "DP02_0063PE")], na.rm = TRUE)) %>%
  rename(state = NAME, tot_high_school_grad = "DP02_0061E", pct_high_school_grad = "DP02_0061PE",
         tot_bachelor = "DP02_0064E", pct_bachelor = "DP02_0064PE",
         tot_grad_prof_degree ="DP02_0065E", pct_grad_prof_degree = "DP02_0065PE") %>%
    slice(-c(1))

# delete variables we don't need 
educ <- educ[,-c(2:5, 8:11)]

# rearrange data.frame
educ <- educ[,c(1,8,2,10,4,6,9,3,11,5,7)]
View(educ)
```


Age & Race, 2013-2018

```{r setup, include=FALSE}

factor_age_race <- read.csv("Age_Race_ACS5_2018.csv", header=TRUE)
View(factor_age_race)
str(factor_age_race)

# Change variables from factor to numeric

my_function <-function(x) {
  as.numeric(levels(x))[x]}

age_race <- factor_age_race %>%
  mutate(NAME = as.character(NAME), GEO_ID=as.character(GEO_ID)) %>%
  mutate_if(is.factor, my_function )
str(age_race)
View(age_race)


## Data Inspection
# # use a vector of first row to use grep
# 
# M <- as.matrix(factor_age_race)
# Row1 <- M[1,]
# str(Row1)
# 
# 
# ### Age Range Find Columns that we need ###
# 
# Row1[grep("Under 5 years", Row1)]
# Row1[grep("5 to 9 years", Row1)]
# Row1[grep("10 to 14 years", Row1)]
# # ...
# # ...
# Row1[grep("60 to 64 years", Row1)]
# 
# grep("Under 5 years", Row1)
# grep("60 to 64 years", Row1)
# # 252-291
# Row1[grep("65 years and over", Row1)]
# # DP05_0024E total number 
# # DP05_0024PE percent 
# 
# ## Create Age Range Variables ## 
# 
# # 
# # # use subsection of factor data.frame to get a better overview of the age data
# # age_range <- factor_age_race %>%
# # select(c(252:291))
# # 
# # View(age_range)
# # 
# # # delete margin errors - delete every second row
# # age_range <- age_range[-c(seq(2,40,2))]
# # 
# # View(age_range)


# Create columns "under 25", "25 to 44", "45 to 65"

age_race<- age_race %>%
  mutate( total_under_25 = rowSums(age_race[, c("DP05_0005E", "DP05_0006E", "DP05_0007E", "DP05_0008E", "DP05_0009E")], na.rm = TRUE)) %>%
  mutate( pct_under_25=  rowSums(age_race[, c("DP05_0005PE", "DP05_0006PE", "DP05_0007PE", "DP05_0008PE", "DP05_0009PE")], na.rm = TRUE)) %>%
  mutate( total_25_44 = rowSums(age_race[, c("DP05_0010E", "DP05_0011E")], na.rm = TRUE)) %>%
  mutate( pct_25_44 =  rowSums(age_race[, c("DP05_0010PE", "DP05_0011PE")], na.rm = TRUE)) %>%
  mutate( total_45_64 = rowSums(age_race[, c("DP05_0012E", "DP05_0013E", "DP05_0014E")], na.rm = TRUE)) %>%
  mutate( pct_45_64 =  rowSums(age_race[, c("DP05_0012PE", "DP05_0013PE", "DP05_0014PE")], na.rm = TRUE))


View(age_race)

## Dta Inspection
# ### RACE Find Columns ### 
# 
# Row1[grep("Total population", Row1)]
# #DP05_0033E
#   
# Row1[grep("Hispanic or Latino", Row1)]
# # DP05_0071E Total Number of Hispanic or Latino of any race
# # DP05_0071PE Percent Hispanic or Latino of any race
# 
# 
# Row1[grep("Chinese", Row1)]
# # DP05_0046E Total Number of Chinese (one race) 
# # DP05_0046PE Percent of Chinese (one race) 
# 
# 
# Row1[grep("Asian", Row1)]
# # DP05_0067E Total Number of Asians (one or more races) 
# # DP05_0067PE Percent of Asians (one or more races) 


age_race <- age_race %>%
  select(NAME, DP05_0033E, DP05_0071E, DP05_0071PE, DP05_0046E, DP05_0046PE, DP05_0067E, DP05_0067PE, total_under_25, pct_under_25, total_25_44,  pct_25_44, total_45_64, pct_45_64, DP05_0024E, DP05_0024PE) %>%
  rename(state=NAME, total_pop = DP05_0033E, total_hispanic = DP05_0071E, percent_hispanic=DP05_0071PE, total_chinese = DP05_0046E, percent_chinese=DP05_0046PE, total_asian = DP05_0067E, percent_asian = DP05_0067PE, total_65andover = DP05_0024E, pct_65andover = DP05_0024PE)%>%
  slice(-c(1))

View(age_race)

# reorder data.frame
age_race <- age_race[,c(1:9,11,13,15,10,12,14,16)]
View(age_race)



# 
# 
# ## not sure if we want that right now, ?? WOULD IT AFFECT OTHER VARIABLES?  
# ## maybe we would need this for the graphics later to be able to use aesthetics like colour ect, but not really neccessary since it's only a minor variable and we will probably use it for plotting   
# #combine age in one column - gather() - longformat
# 
# age_cat_data <- age_race %>%
#   gather(key="age_cat_pct", value="age_pct", 13:16 ) %>%
#   gather(key="age_cat_nr", value="age_nr", 9:12)
# 
# 
# View(age_cat_data)




```


Combine datasets into one sociodemographic dataset

```{r setup, include=FALSE}
help(full_join)
# age_race, educ, empl_HHincome, poverty

sociodemographics <- full_join(age_race, educ, by = "state")
sociodemographics <- full_join(sociodemographics, empl_HHincome, by = "state")
sociodemographics <- full_join(sociodemographics, poverty, by = "state")

View(sociodemographics)
```



Voting result of elections 2016


```{r}
library(xml2)
library(rvest)
```


```{r}
# read page as an R object

src <- read_html("https://en.wikipedia.org/wiki/2016_United_States_presidential_election#Statistical_analysis")

```

Now extract the tables from this object (using the `rvest` package) and save the result as a new object. Follow the instructions if there is an error.

```{r}
# store all tables in object "all_tables"
tables <- html_nodes(src, xpath = '//table')
tables

# convert html tables in data.frames

vote <- html_table(tables[41], fill = TRUE)
summary(vote)
str(vote)

# grep("Donald TrumpRepublican", vote)
# [[1]]
vote <- vote[[1]]

View(vote)

# deleting columns I don't need
vote <- vote[-c(1,58:59),c(1, 8)]

# naming variables
names(vote) <- c("state", "election16_republicans")

# recoding election win (2016) into a dummy variable 
# 1 = republicans won
# 0 = democrats won

vote <- vote %>%
  mutate(election16_republicans = case_when(election16_republicans >= 1 ~ 1,
                                            election16_republicans == '–' ~ 0))
  
# All districts in Nebraska voted for Trump --> use as one variable
vote <- vote[-c(31:33),]

# In Maine the results were different in seperate districts --> assign missing value

vote$election16_republicans[20] <- NA
vote <- vote[-c(21,22),]

## delete "(at-lg)"

# make a matrix
vote <- vote %>% 
  as.matrix()

# delete "(at-lg)" to join data.frame with sociodemograpics data.frame

vote <- gsub("..(at-lg).", "", vote )

vote <- as.data.frame(vote)

class(vote)
str(vote)

View(vote)
```


Join dataset sociodemograpic and vote together 
```{r}

sociodemographics <- full_join(sociodemographics, vote, by = "state")

View(sociodemographics)
str(sociodemographics)

```